Efficient, Robust and Exact Boolean Operations on N Triangular Mesh Primitives

### 2016-09-23 summary.tex
9.23

We use Campen et al.'s method \cite{campen2010exact} for efficient and exact conversion from the vertex-based representation to plane-based representation. Given the input vertex coordinates represented by $L$ bits including sign, the precisions of the four plane coefficients required for exact conversion are $L_a, L_b, L_c, L_d$, where $L_a=L_b=L_c=2(K-1)+1+1$ and $L_d=(L_a-1)+(L-1)+2+1$. Here $K$ is the maximum bits to represent edge vector components. If each coefficient is stored using $M$ bits, the constraint is $M \ge L_d$. In addition, different from Campen et al.'s method, we further enforce the constraint to $M \ge L_d+1$. In this way, given a $L$-bit vertex $\bm{v}$, we can compute predicate of orientation of plane with respect to the vertex by simply compute the sign $\bm{n}\cdot\bm{v} + d$ within $M$-bit precision. This allows fast predicates especially for early rejection in collision detection ($\S$ \ref{section:isect}).

Let $\delta = 2^{K-L-1}$ be the relative length (in max-norm) of the longest edge in the mesh (relative to the bounding box). If we use IEEE 754 double float point number ($M=53$) and assume $\delta \le 2^{-5.5} \approx 0.22$, the maximum input precision $L$ can reach to 20, which is enough for most application (note that IEEE 754 single float point number has only 24-bit precision).

\subsection{Substrates}
\label{sec:substrates}
Plane-based geometry predicates include checking coincidence and co-orientation of planes, orientation of a plane with respect to a point, and whether three planes intersect in a unique point, etc. Most of them are well-discussed in previous work \cite{bernstein2009fast,banerjee1996topologically}. In the following, we focus on two advanced plane-based predicates---\emph{point ordering on line} and \emph{circular ordering of line}.

\vspace{0.5em}
\noindent \textbf{Point ordering on line}~~~~
Given an arbitrary line $l\colon(p_m, p_n)$ with two points $v_p$ and $v_q$ on $l$, the two points can be represented by plane triples $(p_m, p_n, p_p)$ and $(p_m, p_n, p_q)$, sharing the same first two components. We want to determine if $v_q$ is in positive side of $v_p$ along $\vec{l}$. We call this predicate as \emph{two-point-one-line comparison} (Fig. \ref{fig:twopointoneline}). This problem can be converted to the problem of orientation of a plane $p_p$ with respect to a point $v_q$, except that there is an extra request---we have to guarantee $\vec{l}$ and $\vec{n}_p$ have the same orientation. If not, we negate the normal of $p_p$, and the implicit coordinate of  $v_p\colon(p_m, p_n, p_p)$ remains unchanged.

\vspace{0.5em}
\noindent \textbf{Circular ordering of line}~~~~
As in Fig. \ref{fig:substrates}, Given two coplanar but no parallel lines $l_a\colon(p_0, p_a)$ and $l_b\colon(p_0, p_b)$, they intersect on point $\bm{v}\colon(p_0,p_a,p_b)$.we want decide whether the cross product of $\vec{l}_a$ and $\vec{l}_b$ has the same orientation as $\vec{n}_0$ (the normal of $p_0$). Direct solution requires to explicitly compute the vector of $l_a$ and $l_b$ and then decide the sign of $\vec{n}_0 \cdot (\vec{l}_a \times \vec{l}_b)$. However, explicit representation of $\vec{l}_a$ and $\vec{l}_b$ cannot be represented by the float point number.

>>>>

We use Campen et al.'s method \cite{campen2010exact} for exact conversion of triangles (represented by three vertices) from vertex-based representation to plane-based representation. Given that the input vertex coordinate has $L$-bit precision including sign, the precisions of the four plane coefficients are $L_a, L_b, L_c, L_d$, which have to meet the following requirements:
\begin{equation}
\begin{split}
&L_a=L_b=L_c\ge 2(K-1)+1+1,\\
&L_d\ge(L_a-1)+(L-1)+2+1.
\end{split}
\end{equation}
Here $K$ is the maximum bits required to represent edge vector components in input meshes. If we use $M$ bits to store each plane coefficient, basically $M$ has to be greater than $L_d$. In our implementation, we actually use a more strong constraint $M \ge L_d+1$. Under this constraint, given a $L$-bit vertex $\bm{v}$, we can compute the orientation of plane with respect to $\bm{v}$ by the sign $\bm{n}\cdot\bm{v} + d$, which can be exactly computed using common float point arithmetic with limited $M$-bit precision. It allows very fast plane orientation predicates which benefits much the early rejection in our intersection computation ($\S$ \ref{section:isect}).

Let $\delta = 2^{K-L-1}$ be the relative length (in max-norm) of the longest edge in the mesh (relative to the bounding box). If we use IEEE 754 double float point number ($M=53$) and assume $\delta \le 2^{-5.5} \approx 0.022$, the maximum input precision $L$ can reach to 20, which is enough for most application (for standard IEEE 754 single float point number, the precision is 24-bit).

\subsection{Substrates}
\label{sec:substrates}
Plane-based geometry predicates include checking coincidence and co-orientation of planes, orientation of a plane with respect to a point, and whether three planes intersect in a unique point, etc. Most of them are well-discussed in previous work \cite{bernstein2009fast,banerjee1996topologically}. In the following, we focus on two advanced plane-based predicates---\emph{point ordering on line} and \emph{circular ordering of line}, which are important building blocks of our method.

\vspace{0.5em}
\noindent \textbf{Point ordering on line}~~~~
Given an arbitrary line $\bm{l}\colon(p_m, p_n)$ with two points $\bm{v}_1\colon(\bm{p}_1^0\cap\bm{p}_1^1\cap\bm{p}_1^2)$ and $\bm{v}_2\colon(\bm{p}_2^0\cap\bm{p}_2^1\cap\bm{p}_2^2)$ on it, we need to determine the linear order of the two points along $\bm{l}$ (Fig. \ref{fig:twopointoneline}). To solve this problem, we choose one plane not parallel with $\bm{l}$ from the plane representation of each point. Then we convert this problem into linear order of planes and solve it using the same technique as Banerjee et al. \cite{banerjee1996topologically}. The chosen planes have to have the same orientation with $\bm{l}$ (the dot product between plane normal and $\bm{l}$ has to be positive) and unqualified planes are flipped.

\vspace{0.5em}
\noindent \textbf{Circular ordering of line}~~~~
During face tessellation, we need to know the which edges are neighboring. While the relationships among original mesh edges are clear, the relations among newly introduced edges, which are generated by intersection between meshes, are totally unknown. Given two coplanar and not parallel lines $\bm{l}_a\colon(\bm{p}_0\cap\bm{p}_a)$ and $l_b\colon(p_0, p_b)$, they intersect on point $\bm{v}\colon(p_0,p_a,p_b)$.we want decide whether the cross product of $\vec{l}_a$ and $\vec{l}_b$ has the same orientation as $\vec{n}_0$ (the normal of $p_0$). Direct solution requires to explicitly compute the vector of $l_a$ and $l_b$ and then decide the sign of $\vec{n}_0 \cdot (\vec{l}_a \times \vec{l}_b)$. However, explicit representation of $\vec{l}_a$ and $\vec{l}_b$ cannot be represented by the float point number.In Fig. \ref{fig:substrates},
